# Extensions {#sec-extensions}

```{r}
#| label: setup
#| message: false
#| warning: false
#| include: false
library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)
library(colorblindr)
library(downlit)
# _common.R ----
source("_common.R")
# use font
showtext::showtext_auto()
# set theme
ggplot2::theme_set(theme_ggp2g(
    base_size = 15))
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o", look = "minimal",
  header = "Caution",
  contents = "This section is still being developed. The contents are subject to change.",
  fold = FALSE
)
library(dplyr)
```

Below are the graphs requiring additional packages/extensions:

::: {style="font-size: 0.95rem;"}
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Graph           | Package                                                                                                                | Code                                              |
+=================+========================================================================================================================+===================================================+
| Waffle charts   | [`ggwaffle`](https://liamgilbey.github.io/ggwaffle/)                                                                   | ```                                               |
|                 |                                                                                                                        | remotes::install_github("liamgilbey/ggwaffle")    |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Pie charts      | [`ggpubr`](https://rpkgs.datanovia.com/ggpubr/)                                                                        | ```                                               |
|                 |                                                                                                                        | remotes::install_github("kassambara/ggpubr")      |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Mosaic Plots    | [`ggmosaic`](https://haleyjeppson.github.io/ggmosaic/)                                                                 | ```                                               |
|                 |                                                                                                                        | remotes::install_github("haleyjeppson/ggmosaic")  |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Treemaps        | [`treemapify`](https://wilkox.org/treemapify/)                                                                         | ```                                               |
|                 |                                                                                                                        | remotes::install_github("wilkox/treemapify")      |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Beeswarm Plots  | [`ggbeeswarm`](https://github.com/eclarke/ggbeeswarm)                                                                  | ```                                               |
|                 |                                                                                                                        | remotes::install_github("eclarke/ggbeeswarm")     |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Ridgeline Plots | [`ggridges`](https://wilkelab.org/ggridges/)                                                                           | ```                                               |
|                 |                                                                                                                        | remotes::install_github("wilkelab/ggridges")      |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Raincloud plots | [`raincloudplots`](https://github.com/jorvlan/raincloudplots) & [`ggdist`](https://mjskay.github.io/ggdist/reference/) | ```                                               |
|                 |                                                                                                                        | remotes::install_github('jorvlan/raincloudplots') |
|                 |                                                                                                                        | remotes::install_github('mjskay/ggdist')          |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Alluvial charts | [`ggalluvial`](https://corybrunson.github.io/ggalluvial/)                                                              | ```                                               |
|                 |                                                                                                                        | remotes::install_github("corybrunson/ggalluvial") |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Parallel Sets   | [`ggforce`](https://ggforce.data-imaginist.com/)                                                                       | ```                                               |
|                 |                                                                                                                        | remotes::install_github("thomasp85/ggforce")      |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Bump charts     | [`ggbump`](https://github.com/davidsjoberg/ggbump)                                                                     | ```                                               |
|                 |                                                                                                                        | remotes::install_github("davidsjoberg/ggbump")    |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
| Stream Plots    | [`ggstream`](https://github.com/davidsjoberg/ggstream)                                                                 | ```                                               |
|                 |                                                                                                                        | remotes::install_github("davidsjoberg/ggstream")  |
|                 |                                                                                                                        | ```                                               |
+-----------------+------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------+
:::
